Draft Report

### 2020-11-18 DraftReport.Rmd
Updated draft report for consistency and adjusted for knitted PDF output

 **Initial reassignment of common data used across models:**
```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lmtest)
library(sandwich)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(patchwork)
library(stargazer)
library(tinytex)
```

@@

**Research Question:** How is the COVID case rate related to the population density and distribution of demographics of a state? How does the effect of population make-up on case rate compare with the effect of policy decisions made in each state?"

@@

```{r}
ggplot(data = df,
  mapping = aes(x= case_rate_100k))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,6000,600)) +
  labs(title =  "Histogram of COVID Case Rate per 100,000", x = "COVID case rate per 100,000", y = 'Count')+
  scale_x_continuous(breaks=seq(0, 6000, 600))
```

@@

# Plot the new distribution in a histogram
histogram_of_pdensity <- df %>%
  ggplot(aes(x = l_population_density)) + 
  geom_histogram(fill = 'skyblue', color = 'grey30', bins = 20) + 
  labs(
    title = 'Figure 2. Distribution of Population Density per Square Miles', 
    x = 'Log 10 of population density per square mile', y = 'Count')

@@

```{r}
df %>% 
  ggplot(aes(l_population_density, case_rate_100k)) + 
  geom_smooth(se = FALSE) +
  geom_point() +   
  labs(
    title = 'COVID Case Rate due to Population Density', 
    x = 'Log 10 of population density per square miles', 
    y = 'COVID case rate per 100,000'
  )
```

@@

```{r}
black_num <- as.numeric(df$black_population_pct)
black_num[is.na(black_num)]<- 0
cor(df$white_population_pct, black_num)
```

@@

```{r}
ggplot(data = df,
  mapping = aes(x= poverty_pct))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(6,21,1)) +
  labs(title =  "Histogram of Poverty Percentage", x = "Percentage living under poverty line(2018)", y = 'Count')+
  scale_x_continuous(breaks=seq(6, 21, 1))
```

@@

```{r}
df %>% 
  ggplot(aes(white_population_pct, case_rate_100k, color = log10(population_density))) + 
  geom_point() + 
  geom_smooth(se=FALSE)+
  labs(
    title = 'Relation of Case Rate Per 100000 to White Population Percentage', 
    x = 'White percentage of total population', 
    y = 'Case rate per 100000',
    color = 'Log 10 of population density'
  )
```

```{r}
df %>% 
  ggplot(aes(x = white_population_pct, y = case_rate_100k, color = poverty_pct)) + 
  geom_point() +
  geom_smooth(se=FALSE)+
  labs(
    title = 'Relation of Case Rate per 100000 to White Population Percentage', 
    x = 'White percentage of total population', 
    y = 'Case rate per 100000',
    color = 'Poverty percentage'
  )
```

```{r}
df %>% 
  ggplot(aes(x = poverty_pct , y = case_rate_100k, color = log10(population_density))) + 
  geom_point() + 
  geom_smooth(se=FALSE)+
  labs(
    title = 'Relation of Case Rate per 100000 to Poverty Percentage', 
    x = 'Poverty Percentage', 
    y = 'Case rate per 100000',
    color = 'Log 10 of population density'
  )
```

```{r}
df %>% 
  ggplot(aes(unemployed_pct, case_rate_100k, color = log10(population_density))) + 
  geom_point() + 
  geom_smooth(se=FALSE)+
  labs(
    title = 'Relation of Case Rate per 100000 to Unemployed Percentage', 
    x = 'Unemployed percentage(2018)', 
    y = 'Case rate per 100000',
    color = 'log 10 of population density'
  )
```

```{r}
df %>% 
  ggplot(aes(senior_pct, case_rate_100k, color = log10(population_density))) + 
  geom_point() +   
  geom_smooth(se=FALSE)+
  labs(
    title = 'Relation of Case Rate per 100000 to Seniors Percentage(65+)', 
    x = 'Seniors percentage', 
    y = 'Case rate per 100000',
    color = 'log 10 of population density'
  )
```

@@

### Model-2 Regression

@@

df <- df %>% 
  mutate(
    mask_public_bool = case_when(
      mask_public == 0 ~ 0, 
      !(mask_public == 0) ~ 1
      )
  )
head(df)
```

@@

#### Model 3-b&c

@@

```{r}
df$business_close_diff =as.numeric(df$business_closed-mean(df$business_closed, na.rm=TRUE))
head(df)
```

@@

```{r warning=FALSE}
df %>% 
  ggplot(aes(x = business_close_diff, y = case_rate_100k, color = business_close_diff)) + 
  geom_point() + 
  geom_smooth(se = FALSE) +
  labs(
    title = 'Case Rate vs Non-Essential Business Close Date difference compared to Average', 
    x = 'Non-Essential Business Close Date difference compared to Average', 
    y = 'Case Rate', 
    color = 'Case Rate'
  )
```

@@

The findings above suggests that there is colinearity between the two variables "mask_pubic_bool" and "business_close_diff". Therefore, we decided to use "mask_public_bool" variable which has higher significance to represent our policy variable. As a result, model3a is selected as the final model for model 3.

@@

Secondly, the process being modeled must be described by a linear conditional expectation function of the variables which are included, in order for the linear regression to be valid. This can be assessed after the fact by visualizing the model over the range of each variable alongside the data, or also by visualizing the residuals over the range of each variable. For the purpose of this assumption and other model or data related assumptions, only the Model 3A will be used as it is the most inclusive.

@@

Finally, perfect colinearity is automatically detected by the R regression algorithms, which drop potential variables if they have complete colinearity with another variable. Near perfect colinearity is more problematic to detect, but a common hint to its existence is large standard errors on colinear features. In Model 3A, all of the standard errors for model coefficients are at least less than half of the value of the coefficient itself. Therefore, near-perfect colinearity can be assumed to not be taking place between these variables. Conceptually, the variables which may have an important colinearity in the final model are variables such as different percentage make ups of race and ethnicity, or related public policy. These variables are appropriately evaluated and excluded throughout the model building process.

@@

For the regression table below, model 1, model 2b, and model 3a were selected as the most significant models to include.

@@

Thirdly, we notice that the R2 value increases by a large value for every additional variable that is included within the model. Initially, we start with a very low R2 value, suggesting that almost no variance is explained in model 1. However, although still low, there is close to a 9 times increase in the amount of variance that is explained in model 2. Finally, in model 3, the R2 value increases by roughly 2.75 times the previous value, and a decent amount of the variance can be seen to be explained by the model. The adjusted R2 that accounts for the increase in number of variable terms in the model also correspondingly increases as the covariates are added.

@@

Conversely, the hypotheses that population and demographics would have an effect on the COVID case rate was unable to be accepted. Although upon investigating the poverty line percentage variable in model 2 there was a detection of notable adjacency to a significant result (and an actual significant result when using the `r stargazer` to compute the regression table) with a coefficient value of 0.06286, it was not enough to be truly significant and reject the null hypothesis. Similarly, the relationship of race/ethnicity as well as age on case rate was also insignificant with no apparent correlation.

>>>>

 **Initial reassignment of common data used across models:**
```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lmtest)
library(sandwich)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(patchwork)
library(stargazer)
library(tinytex)
library(dplyr)
```

@@

**Research Question:** How is the COVID case rate related to the population density and distribution of demographics of a state? How does the effect of population make-up on case rate compare with the effect of policy decisions made in each state?

@@

```{r}
ggplot(data = df,
  mapping = aes(x= case_rate_100k)) +
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,6000,600)) +
  labs(title =  "Histogram of COVID Case Rate per 100,000", 
       x = "COVID case rate per 100,000", y = 'Count') +
  scale_x_continuous(breaks=seq(0, 6000, 600))
```

@@

# Plot the new distribution in a histogram
histogram_of_pdensity <- df %>%
  ggplot(aes(x = l_population_density)) + 
  geom_histogram(fill = 'skyblue', color = 'grey30', bins = 20) + 
  labs(
    title = 'Distribution of Population Density per Square Miles', 
    x = 'Log of population density per square mile', y = 'Count')

@@

```{r message=FALSE}
df %>% 
  ggplot(aes(l_population_density, case_rate_100k)) + 
  geom_smooth(se = FALSE) +
  geom_point() +   
  labs(
    title = 'COVID Case Rate due to Population Density', 
    x = 'Log of population density per square miles', 
    y = 'COVID case rate per 100,000'
  )
```

@@

```{r warning=FALSE}
black_num <- as.numeric(df$black_population_pct)
black_num[is.na(black_num)]<- 0
cor(df$white_population_pct, black_num)
```

@@

```{r}
ggplot(data = df,
  mapping = aes(x= poverty_pct))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(6,21,1)) +
  labs(title =  "Histogram of Poverty Percentage", 
       x = "Percentage living under poverty line (2018)", y = 'Count')+
  scale_x_continuous(breaks=seq(6, 21, 1))
```

@@

```{r message=FALSE}
df %>% 
  ggplot(aes(white_population_pct, case_rate_100k, color = log10(population_density))) + 
  geom_point() + 
  geom_smooth(se=FALSE)+
  labs(
    title = 'Relation of Case Rate Per 100000 to White Population Percentage', 
    x = 'White percentage of total population', 
    y = 'Case rate per 100000',
    color = 'Log 10 of population density'
  )
```

```{r message=FALSE}
df %>% 
  ggplot(aes(x = white_population_pct, y = case_rate_100k, color = poverty_pct)) + 
  geom_point() +
  geom_smooth(se=FALSE)+
  labs(
    title = 'Relation of Case Rate per 100000 to White Population Percentage', 
    x = 'White percentage of total population', 
    y = 'Case rate per 100000',
    color = 'Poverty percentage'
  )
```

```{r message=FALSE}
df %>% 
  ggplot(aes(x = poverty_pct , y = case_rate_100k, color = log10(population_density))) + 
  geom_point() + 
  geom_smooth(se=FALSE)+
  labs(
    title = 'Relation of Case Rate per 100000 to Poverty Percentage', 
    x = 'Poverty Percentage', 
    y = 'Case rate per 100000',
    color = 'Log 10 of population density'
  )
```

```{r message=FALSE}
df %>% 
  ggplot(aes(unemployed_pct, case_rate_100k, color = log10(population_density))) + 
  geom_point() + 
  geom_smooth(se=FALSE)+
  labs(
    title = 'Relation of Case Rate per 100000 to Unemployed Percentage', 
    x = 'Unemployed percentage (2018)', 
    y = 'Case rate per 100000',
    color = 'log 10 of population density'
  )
```

```{r message=FALSE}
df %>% 
  ggplot(aes(senior_pct, case_rate_100k, color = log10(population_density))) + 
  geom_point() +   
  geom_smooth(se=FALSE)+
  labs(
    title = 'Relation of Case Rate per 100000 to Seniors Percentage (65+)', 
    x = 'Seniors percentage', 
    y = 'Case rate per 100000',
    color = 'log 10 of population density'
  )
```

@@

### Model 2 Regression

@@

df <- df %>% 
  mutate(
    mask_public_bool = case_when(
      mask_public == 0 ~ 0, 
      !(mask_public == 0) ~ 1
      )
  )
```

@@

#### Model 3-b & 3-c

@@

```{r}
df$business_close_diff =as.numeric(df$business_closed-mean(df$business_closed, na.rm=TRUE))
```

@@

```{r warning=FALSE, message=FALSE}
df %>% 
  ggplot(aes(x = business_close_diff, y = case_rate_100k, color = business_close_diff)) + 
  geom_point() + 
  geom_smooth(se = FALSE) +
  labs(
    title = 'Case Rate vs Non-Essential Business Close Date difference \ncompared to Average', 
    x = 'Non-Essential Business Close Date difference compared to Average', 
    y = 'Case Rate', 
    color = 'Case Rate'
  )
```

@@

The findings above suggests that there is colinearity between the two variables "mask_pubic_bool" and "business_close_diff". Therefore, we decided to use "mask_public_bool" variable which has higher significance to represent our policy variable. As a result, model 3-a is selected as the final model for model 3.

@@

Secondly, the process being modeled must be described by a linear conditional expectation function of the variables which are included, in order for the linear regression to be valid. This can be assessed after the fact by visualizing the model over the range of each variable alongside the data, or also by visualizing the residuals over the range of each variable. For the purpose of this assumption and other model or data related assumptions, only the Model 3-a will be used as it is the most inclusive.

@@

Finally, perfect colinearity is automatically detected by the R regression algorithms, which drop potential variables if they have complete colinearity with another variable. Near perfect colinearity is more problematic to detect, but a common hint to its existence is large standard errors on colinear features. In Model 3-a, all of the standard errors for model coefficients are at least less than half of the value of the coefficient itself. Therefore, near-perfect colinearity can be assumed to not be taking place between these variables. Conceptually, the variables which may have an important colinearity in the final model are variables such as different percentage make ups of race and ethnicity, or related public policy. These variables are appropriately evaluated and excluded throughout the model building process.

@@

For the regression table below, model 1, model 2-b, and model 3-a were selected as the most significant models to include.

@@

Thirdly, we notice that the R2 value increases by a large value for every additional variable that is included within the model. Initially, we start with a very low R2 value, suggesting that almost no variance is explained in model 1 and that the model is a poor fit and predictor for the data. However, although still low, there is close to a 9 times increase in the amount of variance that is explained in model 2. Finally, in model 3, the R2 value increases by roughly 2.75 times the previous value, and a decent amount of the variance can be seen to be explained by the model. The adjusted R2 that accounts for the increase in number of variable terms in the model also correspondingly increases as the covariates are added. It also holds a negative value for model 1, signifying that there is no variance that can be explained by the model, further highlighting the poor fit.

@@

Conversely, the hypotheses that population and demographics would have an effect on the COVID case rate was unable to be accepted. Although upon investigating the poverty line percentage variable in model 2 there was a detection of notable adjacency to a significant result (and an actual significant result when using the `stargazer` package to compute the regression table) with a coefficient value of 0.06286, it was not enough to be truly significant and reject the null hypothesis. Similarly, the relationship of race/ethnicity as well as age on case rate was also insignificant with no apparent correlation.
